---
title: "Descriptive Statistics"
author: "Annabel Critchlow"
date: "2024-11-14"
output:
  html_document:
    theme: readable
    toc: TRUE
---


# Load libraries

```{r}

library(tidyverse)
library(readxl)

```

# Load files

```{r}
data <- read_excel("FAMe_Data.xlsx",
                   na = c("","NA", "?", "#VALUE!"),
                 trim_ws = TRUE)

immuno_data <- data %>%
  filter(ihc == "yes")

rna_data <- data %>%
  filter(!is.na(rnaseq) & rnaseq != "", id != "92")
    
```


# Define variables

Variables have been allocated into numerical or categorical.

```{r}
numvars <- c("weight", "height", "bmi", "bodyfat", "TBFM", "TBLM", "ALM", "ALMh", "bmc", "thighmuscle", "intrafat", "subfat", "permuscle", "bonedensity", "onerep", "onerepllm", "onerepcsa", "E2", "FEI", "prog", "TT", "FAI", "TEratio", "protein", "mvpa")

catvars <- c("meno", "hc", "hrt")

immunonumvars <- c("weight", "height", "bmi", "bodyfat", "TBFM", "TBLM", "ALM", "ALMh", "bmc", "thighmuscle", "intrafat", "subfat", "permuscle", "bonedensity", "Ipernumber", "Iperarea", "Imeanarea", "IIApernumber", "IIAperarea", "IIAmeanarea", "Hybridpernumber", "Hybridperarea", "Hybridmeanarea", "Tmeanarea", "onerep", "onerepllm", "onerepcsa", "E2", "FEI", "prog", "TT", "FAI", "TEratio", "protein", "mvpa")

```


# Participant characteristics (Table 2)

This code calculates the mean and standard deviation for all numeric variables, and the frequency of categorical variables, both in the whole cohort and across each age group. Replace 'data' with the data subset (immuno_data or rna_data) to calculate values for the cohort subsets. Replace numvars with immunonumvars when analysing IHC data subset.

```{r}

# Step 1: Numeric summary by agecat
by_agecat_num <- data %>%
  group_by(agecat) %>%
  summarise(across(all_of(numvars),
                   list(mean = ~round(mean(.x, na.rm = TRUE), 1),
                        sd   = ~round(sd(.x, na.rm = TRUE), 1)),
                   .names = "{.col}_{.fn}")) %>%
  pivot_longer(-agecat, names_to = c("Variable", "Stat"), names_sep = "_") %>%
  pivot_wider(names_from = Stat, values_from = value) %>%
  mutate(Value = paste0(mean, " (", sd, ")")) %>%
  select(agecat, Variable, Value)

# Step 2: Overall numeric summary
overall_num <- data %>%
  summarise(across(all_of(numvars),
                   list(mean = ~round(mean(.x, na.rm = TRUE), 1),
                        sd   = ~round(sd(.x, na.rm = TRUE), 1)),
                   .names = "{.col}_{.fn}")) %>%
  pivot_longer(everything(), names_to = c("Variable", "Stat"), names_sep = "_") %>%
  pivot_wider(names_from = Stat, values_from = value) %>%
  mutate(Value = paste0(mean, " (", sd, ")"),
         agecat = "Overall") %>%
  select(agecat, Variable, Value)

# Step 3: Combine numeric summary
summary_num <- bind_rows(by_agecat_num, overall_num) %>%
  pivot_wider(names_from = agecat, values_from = Value) %>%
  mutate(Stat = "mean (sd)") %>%
  select(Variable, Stat, everything())

# Step 4: Count total per group (agecat) for percentage calculation
agecat_totals <- data %>%
  group_by(agecat) %>%
  summarise(N = n(), .groups = "drop")

# Step 5: Categorical summary by agecat as "n / total (percent%)"
by_agecat_cat <- rna_data %>%
  select(agecat, all_of(catvars)) %>%
  pivot_longer(-agecat, names_to = "Variable", values_to = "Level") %>%
  group_by(agecat, Variable, Level) %>%
  summarise(n = n(), .groups = "drop") %>%
  left_join(agecat_totals, by = "agecat") %>%
  mutate(pct = round(100 * n / N, 1),
         Value = paste0(n, " / ", N, " (", pct, "%)")) %>%
  unite(Variable, Variable, Level, sep = ": ") %>%
  select(agecat, Variable, Value) %>%
  pivot_wider(names_from = agecat, values_from = Value)

# Step 6: Overall counts for categorical variables
overall_cat <- data %>%
  select(all_of(catvars)) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Level") %>%
  group_by(Variable, Level) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(N = nrow(data),
         pct = round(100 * n / N, 0),
         Overall = paste0(n, " / ", N, " (", pct, "%)")) %>%
  unite(Variable, Variable, Level, sep = ": ") %>%
  select(Variable, Overall)

# Step 7: Combine categorical summaries
summary_cat <- full_join(by_agecat_cat, overall_cat, by = "Variable") %>%
  mutate(Stat = "count / total (%)") %>%
  select(Variable, Stat, everything())

# Step 8: Final combined summary table
final_table <- bind_rows(summary_num, summary_cat)

# Step 9: Export
print(final_table, n = Inf)
write_csv(final_table, "Descriptives/Participant_characteristics.csv")

```
