---
title: "Muscle Mass and Function Analysis"
author: "Annabel Critchlow"
date: "2025-02-11"
output: html_document
---

# Load libraries

```{r}
library(tidyverse)     
library(readxl)        
library(broom)         
library(knitr)         
library(kableExtra)    
library(patchwork)     
library(sandwich)      
library(lmtest)        
library(car)           
library(corrplot)      
library(tibble)  
```

# Hormonal effects (Table 3)

This code tests the relationships between reproductive factors (menstrual cycle phase, hormonal contraception, hormonal replacement therapy) and the outcome variables using a linear model. 

## 1. Load unimputed data

```{r}
unimputed_data <- read_excel("FAMe_Data.xlsx",
                   na = c("","NA", "?", "#VALUE!"),
                 trim_ws = TRUE)
```

## 2a. Menstrual cycle phase

neurophase refers to the participant's menstrual cycle phase during their muscle function visit. 

```{r}
# Create subset of data with only premenopausal participants
mcdata <- unimputed_data %>%
  filter(meno == "pre", neurophase != "unknown", hc != "OCP", hc != "Implant") %>%
  drop_na(neurophase)

# Select which variables to analyse and place them in a list
vars <- mcdata %>%
  dplyr::select("onerep", "onerepllm", "onerepcsa")

vars <- names(vars)

models = lapply(vars, function(x) {
  lm(substitute(i ~ neurophase, list(i = as.name(x))), data = mcdata)
})

y=lapply(models, summary)
names(y) <- vars # To add names of the model instead of list number
df=purrr::map_df(y, broom::tidy, .id = 'vars')%>%
  filter(term == "neurophase" | term != "(Intercept)")

knitr::kable(df) %>%
  kable_styling() %>%
  row_spec(which(df$term == "neurophase"  & df$p.value < 0.05), bold = T)

write.csv(df, "Hormonal_effects/Menstrual_cycle_effects.csv")

```

## 2b. Hormonal contraception

```{r}
# Create subset of data with only premenopausal participants
predata <- unimputed_data %>%
  filter(meno == "pre") %>%
  drop_na(hc) %>%
  mutate(hc = factor(hc, levels = c("none", "ocp", "iud", "implant")))

# Select which variables to analyse and place them in a list
vars <- predata %>%
  dplyr::select("bodyfat", "TBFM", "TBLM", "ALM", "ALMh", "thighmuscle", "intrafat", "subfat", "permuscle", "onerep", "onerepllm", "onerepcsa")

vars <- names(vars)

models = lapply(vars, function(x) {
  lm(substitute(i ~ hc, list(i = as.name(x))), data = predata)
})

y=lapply(models, summary)
names(y) <- vars # To add names of the model instead of list number
df=purrr::map_df(y, broom::tidy, .id = 'vars')%>%
  filter(term == "hc" | term != "(Intercept)")

knitr::kable(df) %>%
  kable_styling() %>%
  row_spec(which(df$term == "hc"  & df$p.value < 0.05), bold = T)

write.csv(df, "Hormonal_effects/Contraceptive_effects.csv")

```

## 2c. Hormonal replacement therapy

```{r}
# Create subset of data with only perimenopausal and postmenopausal participants
peripostdata <- unimputed_data %>%
  filter(meno != "pre") %>%
  drop_na(hrt) %>%
  mutate(hrt = factor(hrt, levels = c("no", "yes")))

# Select which variables to analyse and place them in a list
vars <- peripostdata %>%
  dplyr::select("bodyfat", "TBFM", "TBLM", "ALM", "ALMh", "thighmuscle", "intrafat", "subfat", "permuscle", "onerep", "onerepllm", "onerepcsa")

vars <- names(vars)

models = lapply(vars, function(x) {
  lm(substitute(i ~ hrt, list(i = as.name(x))), data = peripostdata)
})

y=lapply(models, summary)
names(y) <- vars
df=purrr::map_df(y, broom::tidy, .id = 'vars')%>%
  filter(term == "hrt" | term != "(Intercept)")

knitr::kable(df) %>%
  kable_styling() %>%
  row_spec(which(df$term == "hrt"  & df$p.value < 0.05), bold = T)

write.csv(df, "Hormonal_effects/HRT_effects.csv")
```

# Age associations (Figure 2)

## 1. Load imputed data

```{r}
data <- read_excel("FAMe_Imputed_Data.xlsx",
                   na = c("","NA", "?", "#VALUE!"),
                 trim_ws = TRUE)
```

## 2. Define variables

```{r}
allvariables <-c("bodyfat", "ALM", "ALMh", "thighmuscle", "intrafat", "subfat", "permuscle", "onerep", "onerepllm", "onerepcsa")

var_labels <-c(
  bodyfat = expression("Body fat (%)"),
  ALM = expression("ALM (kg)"),
  ALMh = expression(bold("ALM index (kg.m"^-2*")")),  
  thighmuscle = expression(bold("Thigh muscle CSA (cm"^2*")")),
  intrafat = expression(bold("Intramuscular fat (cm"^2*")")),  
  subfat = expression(bold("Subcutaneous fat (cm"^2*")")), 
  permuscle = expression("Thigh muscle (%)"),
  onerep = expression("e1RM (kg)"),
  onerepllm = expression(bold("Specific e1RM (kg.kg"^-2*")")),
  onerepcsa = expression(bold("Specific e1RM (kg.cm"^-2*")"))
)

```

## 3. Testing assumptions

Models have been adjusted for protein intake (protein) and physical activity (mvpa)

1. Compare spread of residuals and normality in linear vs. polynomial models
2. Use AIC test to compare score between linear and polynomial (lower score = better)

### 3a. Residuals

```{r}
# Linear Residual plots

# Initialize a list to store the combined plots for each variable
linear_residual_plots <- list()

# Loop over each outcome variable
for (outcome in allvariables) {
  
  # Fit a linear model
  lm_model <- lm(as.formula(paste(outcome, "~ age + protein + mvpa")), data = data)
  
  # Extract residuals and fitted values
  residuals <- resid(lm_model)
  fitted <- fitted(lm_model)
  
  # Create a data frame for plotting
  plot_data <- data.frame(fitted = fitted, residuals = residuals)
  
  # Residuals vs Fitted plot
  p1 <- ggplot(plot_data, aes(x = fitted, y = residuals)) +
    geom_point(alpha = 0.6) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    ggtitle(paste("Residuals vs Fitted for", outcome)) +
    theme_minimal()
  
  # Q-Q Plot of residuals
  p2 <- ggplot(plot_data, aes(sample = residuals)) +
    stat_qq(alpha = 0.6) +
    stat_qq_line() +
    ggtitle(paste("Q-Q Plot of Residuals for", outcome)) +
    theme_minimal()
  
  # Histogram of residuals
  p3 <- ggplot(plot_data, aes(x = residuals)) +
    geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
    ggtitle(paste("Histogram of Residuals for", outcome)) +
    theme_minimal()
  
  # Combine the three plots vertically
  combined_plot <- p1 / p2 / p3
  
  # Add the combined plot to the list
  linear_residual_plots[[outcome]] <- combined_plot
}

# Save all combined plots to a PDF, one page per variable
pdf("Age_associations/All/Linear_Model_Assumptions.pdf", width = 8, height = 11)

for (plot in linear_residual_plots) {
  print(plot)
}

dev.off()

# Polynomial residual plots

# Initialize a list to store the combined plots for each variable
polynomial_residual_plots <- list()

# Loop over each outcome variable
for (outcome in allvariables) {
  
  # Fit a linear model
  lm_model <- lm(as.formula(paste(outcome, "~ poly(age, 2) + protein + mvpa")), data = data)
  
  # Extract residuals and fitted values
  residuals <- resid(lm_model)
  fitted <- fitted(lm_model)
  
  # Create a data frame for plotting
  plot_data <- data.frame(fitted = fitted, residuals = residuals)
  
  # Residuals vs Fitted plot
  p1 <- ggplot(plot_data, aes(x = fitted, y = residuals)) +
    geom_point(alpha = 0.6) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    ggtitle(paste("Residuals vs Fitted for", outcome)) +
    theme_minimal()
  
  # Q-Q Plot of residuals
  p2 <- ggplot(plot_data, aes(sample = residuals)) +
    stat_qq(alpha = 0.6) +
    stat_qq_line() +
    ggtitle(paste("Q-Q Plot of Residuals for", outcome)) +
    theme_minimal()
  
  # Histogram of residuals
  p3 <- ggplot(plot_data, aes(x = residuals)) +
    geom_histogram(bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
    ggtitle(paste("Histogram of Residuals for", outcome)) +
    theme_minimal()
  
  # Combine the three plots vertically
  combined_plot <- p1 / p2 / p3
  
  # Add the combined plot to the list
  polynomial_residual_plots[[outcome]] <- combined_plot
}

# Save all combined plots to a PDF, one page per variable
pdf("Age_associations/All/Polynomial_Model_Assumptions.pdf", width = 8, height = 11)

for (plot in polynomial_residual_plots) {
  print(plot)
}

dev.off()


```

### 3b. AIC

```{r}
# Testing the fit of models with AIC
aic_results <- data.frame(
  Variable = character(),
  Linear_AIC = numeric(),
  Polynomial_AIC = numeric(),
  stringsAsFactors = FALSE
)

# Loop over each variable
for (outcome in allvariables) {
  # Construct formulas
  linear_formula <- as.formula(paste(outcome, "~ age + mvpa + protein"))
  polynomial_formula <- as.formula(paste(outcome, "~ poly(age, 2) + mvpa + protein"))
  
  # Fit linear model
  lm_linear <- lm(linear_formula, data = data)
  
  # Fit polynomial (quadratic) model
  lm_poly <- lm(polynomial_formula, data = data)
  
  # Extract AICs
  aic_linear <- AIC(lm_linear)
  aic_poly <- AIC(lm_poly)
  
  # Append results to the data frame
  aic_results <- rbind(
    aic_results,
    data.frame(Variable = outcome, Linear_AIC = aic_linear, Polynomial_AIC = aic_poly)
  )
}

# Write the results to a CSV file
write.csv(aic_results, "Age_associations/All/Adjusted_Linear_Poly_Model_fit.csv", row.names = FALSE)

```

## 4. Results

This code runs linear regression between age and outcome variables, while adjusting for mvpa and protein.

```{r}

# Create a data frame to collect all model results
all_model_results <- data.frame()

# Loop over each outcome variable
for (outcome in allvariables) {
  
  # Create the formula for the linear model (outcome vs. age + mvpa)
  formula_model <- as.formula(paste(outcome, "~ age + mvpa + protein"))
  
  # Fit the linear model using your data
  model <- lm(formula_model, data = data)
  
  # Get the summary of the model
  model_summary <- summary(model)
  
  # Extract model coefficients and statistics
  model_coefficients <- data.frame(
    Term        = rownames(coef(summary(model))),
    Coefficient = coef(model),
    Std_Error   = coef(summary(model))[ , "Std. Error"],
    t_value     = coef(summary(model))[ , "t value"],
    p_value     = coef(summary(model))[ , "Pr(>|t|)"]
  )
  
  # Confidence intervals
  conf_intervals <- confint(model)
  model_coefficients$CI_Lower <- conf_intervals[ , 1]
  model_coefficients$CI_Upper <- conf_intervals[ , 2]
  
  # R-squared
  R_squared <- model_summary$r.squared
  
  # Label columns
  model_coefficients$Model   = "Adjusted"
  model_coefficients$Outcome = outcome
  model_coefficients$R_squared = R_squared
  
  # Reorder columns
  model_coefficients <- model_coefficients[ , c("Model", "Outcome", "Term", "Coefficient",
                                              "Std_Error", "CI_Lower", "CI_Upper",
                                              "p_value", "R_squared")]
  # Append to the results data frame
  all_model_results <- rbind(all_model_results, model_coefficients)
  
}  # End outcome loop

# Save the combined model results to a CSV file
write.csv(all_model_results, file = "Age_associations/All/Linear_results.csv", row.names = FALSE)

```

## 5. Plots

This code produces plots representing the association between age and each outcome variable. The slope of the line represents the coefficient after adjusting for protein and mvpa. 

```{r}

# Initialize a list to store plots
plot_list <- list()

# Create a data frame to collect all model results
all_model_results <- data.frame()

# Define the list of "nice" intervals
nice_intervals <- c(0.1, 0.2, 0.5, 1, 2, 5, 10, 20, 50, 100, 200, 500, 1000)

# Loop over each outcome variable
for (outcome in allvariables) {
  
  # Create the formula for the linear model (outcome vs. age + mvpa)
  formula_model <- as.formula(paste(outcome, "~ age + mvpa + protein"))
  
  # Fit the linear model using your data
  model <- lm(formula_model, data = data)
  
  # Get the summary of the model
  model_summary <- summary(model)
  
  # Extract the coefficient and p-value for 'age'
  age_coef <- coef(model)["age"]
  age_p_value <- coef(summary(model))["age", "Pr(>|t|)"]
  
  # Format the labels for display
  beta_formatted <- sprintf("%.2f", age_coef)
  p_value_formatted <- sprintf("%.3f", age_p_value)
  if (p_value_formatted == "0.000") {
    p_value_formatted <- "<0.001"
  }
  
  # Define x and y variables
  x_var <- data[["age"]]
  y_var <- data[[outcome]]
  
  # Get the outcome variable label (replace with actual labels if available)
  y_label <- var_labels[[outcome]]  
  x_label <- "Age (years)"
  
  # Prepare the data for plotting (points)
  plot_data <- data.frame(
    x_var           = x_var,
    y_var           = y_var
  )
  
  y_min <- 0
  y_max <- max(y_var, na.rm = TRUE)
  
  raw_interval <- y_max / 8
  interval     <- nice_intervals[which(nice_intervals >= raw_interval)[1]]
  
  y_max_extended <- ceiling(y_max / interval) * interval
  
  y_breaks <- seq(y_min, y_max_extended, by = interval)
  y_limits <- c(y_min, y_max_extended)
  
  # Generate a sequence of age values across its observed range
  age_seq <- seq(
    from = min(x_var, na.rm = TRUE),
    to   = max(x_var, na.rm = TRUE),
    length.out = 100
  )
  
  # Hold covariates at the mean (or median/other typical value)
  mean_mvpa <- mean(data$mvpa, na.rm = TRUE)
  mean_protein <- mean(data$protein, na.rm = TRUE)
  
  pred_data <- expand.grid(
    age            = age_seq,
    mvpa           = mean_mvpa,
    protein        = mean_protein
  )
  
  preds <- predict(model, newdata = pred_data, se.fit = TRUE)
  
  # Add fit + 95% CI to pred_data
  pred_data$fit <- preds$fit
  pred_data$se  <- preds$se.fit
  pred_data$upr <- pred_data$fit + 1.96 * pred_data$se
  pred_data$lwr <- pred_data$fit - 1.96 * pred_data$se
  
  p <- ggplot(plot_data, aes(x = x_var, y = y_var)) +
    # Raw points
    geom_point(aes(), color = "darkblue", size = 1) +
    
    # The adjusted model line in RED (holding mvpa at mean)
    geom_line(
      data = pred_data,
      aes(x = age, y = fit),
      color = "red",         # fixed red line
      size = 1.2,
      inherit.aes = FALSE
    ) +
    
    # Confidence ribbon in RED, partially transparent
    geom_ribbon(
      data = pred_data,
      aes(x = age, ymin = lwr, ymax = upr),
      alpha = 0.2,
      inherit.aes = FALSE
    ) +
    
    # Annotation for Beta and p-value
   annotate(
  "text",
  x = Inf,
  y = if (outcome %in% c("permuscle")) {
    y_min  # position at bottom of plot for permuscle
  } else {
    Inf   # position at top for other outcomes
  },
  label = bquote("B" == .(beta_formatted) ~ "," ~ italic(p) == .(p_value_formatted)),
  hjust = 1.05,
  vjust = if (outcome %in% c("permuscle")) -0.5 else 1.5,
  size = 3,
  color = "black",
  fontface = "bold"
) +
    
    # Labels and scales
    labs(x = x_label, y = y_label)+
    theme(axis.title.y = element_text(face = "bold")) +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 5), expand = c(0, 0.5)) +
    scale_y_continuous(breaks = y_breaks, limits = y_limits, expand = c(0, 0)) +
    
    # Theme
    theme_minimal(base_size = 16) +
    theme(
      text              = element_text(face = "bold", family = "sans"),
      axis.title        = element_text(size = 9, color = "black"),
      axis.line         = element_line(color = "black"),
      axis.text         = element_text(size = 8, color = "black"),
      axis.ticks        = element_line(color = "black", size = 0.8),
      axis.ticks.length = unit(0.2, "cm"),
      panel.grid.major  = element_blank(),
      panel.grid.minor  = element_blank(),
      panel.background  = element_rect(fill = "white", color = "white")
    )
  
  # Save this plot in the list
  plot_list[[length(plot_list) + 1]] <- p
  
  model_coefficients <- data.frame(
    Term        = rownames(coef(summary(model))),
    Coefficient = coef(model),
    Std_Error   = coef(summary(model))[ , "Std. Error"],
    t_value     = coef(summary(model))[ , "t value"],
    p_value     = coef(summary(model))[ , "Pr(>|t|)"]
  )
  
  # Confidence intervals
  conf_intervals <- confint(model)
  model_coefficients$CI_Lower <- conf_intervals[ , 1]
  model_coefficients$CI_Upper <- conf_intervals[ , 2]
  
  # R-squared
  R_squared <- model_summary$r.squared
  
  # Label columns
  model_coefficients$Model   = "Adjusted"
  model_coefficients$Outcome = outcome
  model_coefficients$R_squared = R_squared
  
  # Reorder columns
  model_coefficients <- model_coefficients[ , c("Model", "Outcome", "Term", "Coefficient",
                                                "Std_Error", "CI_Lower", "CI_Upper",
                                                "p_value", "R_squared")]
  # Append to the results data frame
  all_model_results <- rbind(all_model_results, model_coefficients)
  
}  # End outcome loop

combined_plot <- wrap_plots(plot_list, ncol = 4) + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 12))

# Save the combined plot to a PDF file
tiff_filename <- "Age_associations/All/Adjusted_Linear_Plots.tiff"
ggsave(tiff_filename, combined_plot, width = 12, height = 8, dpi = 600, device = "tiff")

```

# Hormone associations (Figure 4)

## 1. Load imputed data

```{r}
data <- read_excel("FAMe_Imputed_Data.xlsx",
                   na = c("","NA", "?", "#VALUE!"),
                 trim_ws = TRUE)
```

## 2. Define variables

Separate lists have been created for body composition and muscle function variables as they have been adjusted for by different variables.

```{r}

# Define outcome variables for hormone association analysis
bodycompvariables <- c("bodyfat", "ALM", "ALMh", "thighmuscle", "intrafat", "subfat", "permuscle")

functionvariables <-c("onerep", "onerepllm", "onerepcsa", "mvc", "rtd", "pt_100", "pt_10", "pt_ratio", "va")

allvariables <-c("bodyfat", "ALM", "ALMh", "thighmuscle", "intrafat", "subfat", "permuscle", "onerep", "onerepllm", "onerepcsa", "mvc", "rtd", "pt_100", "pt_10", "pt_ratio", "va")

hormones <-c(hormonevariables <- c("E2", "FEI", "TT", "FAI", "prog", "TEratio"))

# Mapping variable names to full names
bodycomp_var_labels <- c(
  bodyfat = "Body fat (%)",
  ALM = "Appendicular lean mass (kg)",
  ALMh = expression("ALM index (kg/m"^2*")"),  
  thighmuscle = expression("Thigh muscle CSA (cm"^2*")"),
  intrafat = expression("Intramuscular fat (cm"^2*")"),  
  subfat = expression("Subcutaneous fat (cm"^2*")"),  
  permuscle = "Relative thigh muscle (%)"
)

function_var_labels <-c(
  onerep = "e1RM (kg)",
  onerepllm = "Normalised e1-RM (kg/kg LLM)",
  onerepcsa = expression("Normalised e1RM (kg/cm"^2*")")
)

all_var_labels <- c(
  bodyfat = "Body fat (%)",
  ALM = "Appendicular lean mass (kg)",
  ALMh = expression("ALM index (kg/m"^2*")"),  
  thighmuscle = expression("Thigh muscle CSA (cm"^2*")"),
  intrafat = expression("Intramuscular fat (cm"^2*")"),  
  subfat = expression("Subcutaneous fat (cm"^2*")"), 
  permuscle = "Relative thigh muscle (%)",
  onerep = "e1RM (kg)",
  onerepllm = "Normalised e1RM (kg/kg)",
  onerepcsa = expression("Normalised e1RM (kg/cm"^2*")")
)
```

## 3. Testing assumptions

This code tests the assumptions for using linear regression, including heteroscedasticity, normality, and co-linearity of covariates (using variance inflation factors (VIF)).

Body composition and muscle mass variables were adjusted for age, mvpa, and protein.
Muscle function variables were adjusted for age, mvpa, protein, and bmi.

```{r}
# Define covariates
covariates <- c("age", "mvpa", "protein")

# Initialize empty data frames to store combined results
all_adjusted_results <- data.frame()
all_vif_results <- data.frame()

for (hormone in hormones) {
  # Initialize empty data frames to store results
  adjusted_results <- data.frame()
  vif_results <- data.frame()

  # Create a PDF file to save the adjusted model plots
  pdf(file = paste0("Hormone_associations/All/Bodycomp_Residual_Plots_", hormone, ".pdf"))

  # Loop over each outcome variable for adjusted models
  for (outcome in bodycompvariables) {
    # Create the formula for the adjusted linear model
    formula_adjusted <- as.formula(
      paste(outcome, "~", paste(c(hormone, covariates), collapse = " + "))
    )

    # Fit the linear model
    model_adjusted <- lm(formula_adjusted, data = data)

    # Extract residuals and fitted values
    residuals <- resid(model_adjusted)
    fitted_values <- fitted(model_adjusted)

    # Set up the plotting area to have three plots per page
    par(mfrow = c(3, 1))

    # 1. Residuals vs Fitted Plot
    plot(fitted_values, residuals,
         main = paste("Adjusted Model Residuals vs Fitted for", outcome, "and", hormone),
         xlab = "Fitted Values",
         ylab = "Residuals")
    abline(h = 0, col = "red")

    # 2. Q-Q Plot of Residuals
    qqnorm(residuals,
           main = paste("Adjusted Model Normal Q-Q Plot for", outcome, "and", hormone),
           ylab = "Quantiles of Residuals")
    qqline(residuals, col = "red")

    # 3. Histogram of Residuals
    hist(residuals,
         main = paste("Adjusted Model Histogram of Residuals for", outcome, "and", hormone),
         xlab = "Residuals",
         breaks = 20)

    # Extract model summary statistics using broom
    tidy_adjusted <- tidy(model_adjusted, conf.int = TRUE)
    glance_adjusted <- glance(model_adjusted)

    # Add model and outcome information
    tidy_adjusted$Model <- "Adjusted"
    tidy_adjusted$Outcome <- outcome
    tidy_adjusted$Hormone <- hormone

    # Get R-squared from glance
    tidy_adjusted$R_squared <- glance_adjusted$r.squared

    # Get AIC from glance
    tidy_adjusted$AIC <- glance_adjusted$AIC

    # Reorder columns (now including AIC)
    tidy_adjusted <- tidy_adjusted[, c(
      "Model", "Outcome", "Hormone",
      "term", "estimate", "std.error",
      "conf.low", "conf.high",
      "p.value", "R_squared",
      "AIC"
    )]

    colnames(tidy_adjusted) <- c(
      "Model", "Outcome", "Hormone",
      "Term", "Coefficient", "Std_Error",
      "CI_Lower", "CI_Upper",
      "p_value", "R_squared",
      "AIC"
    )

    # Append to the adjusted results data frame
    adjusted_results <- rbind(adjusted_results, tidy_adjusted)

    # Calculate VIFs for the adjusted model
    vif_values <- vif(model_adjusted)
    vif_df <- data.frame(
      Model = "Adjusted",
      Outcome = outcome,
      Hormone = hormone,
      Term = names(vif_values),
      VIF = vif_values
    )

    # Append to the VIF results data frame
    vif_results <- rbind(vif_results, vif_df)
  }

  # Close the PDF device for adjusted models
  dev.off()

  # Append to the combined adjusted results data frame
  all_adjusted_results <- rbind(all_adjusted_results, adjusted_results)

  # Append to the combined VIF results data frame
  all_vif_results <- rbind(all_vif_results, vif_results)
}

# Save the combined adjusted results to a CSV file
write.csv(
  all_adjusted_results,
  file = "Hormone_associations/All/Function_Adjusted_results_age_mvpa_protein.csv",
  row.names = FALSE
)

# Save the combined VIF results to a CSV file
write.csv(
  all_vif_results,
  file = "Hormone_associations/All/Function_VIF_results_age_mvpa_protein.csv",
  row.names = FALSE
)

```

## 4. Results

This code uses standardised robust models to test the associations between hormones and the outcome variables. 

Body composition and muscle mass variables were adjusted for age, mvpa, and protein.
Muscle function variables were adjusted for age, mvpa, protein, and bmi.

```{r}
covariates <- c("age","mvpa", "protein")

# Standardise the specific variables
data_standardised <- data %>%
  mutate(across(all_of(c(allvariables, covariates, hormones)), scale))

# Initialize empty data frames to store combined results
all_robust_adjusted_results <- data.frame()

# Loop over each hormone
for (hormone in hormones) {
  
  # Initialize empty data frame to store results
  robust_adjusted_results <- data.frame()
  
  # Loop over each outcome variable for adjusted models with robust standard errors
  for (outcome in bodycompvariables) {
    
    # Create the formula for the adjusted linear model
    formula_adjusted <- as.formula(paste(outcome, "~", paste(c(hormone, covariates), collapse = " + ")))
    
    # Fit the linear model using standardized data
    model_adjusted <- lm(formula_adjusted, data = data_standardised)
    
    # Calculate robust standard errors
    robust_se <- vcovHC(model_adjusted, type = "HC1")
    robust_summary <- coeftest(model_adjusted, vcov = robust_se)
    
    # Extract coefficients and p-values from robust summary
    robust_coefficients <- data.frame(
      Term = rownames(robust_summary),
      Coefficient = robust_summary[, "Estimate"],
      Std_Error = robust_summary[, "Std. Error"],
      t_value = robust_summary[, "t value"],
      p_value = robust_summary[, "Pr(>|t|)"]
    )
    
    # Calculate confidence intervals manually
    alpha <- 0.05  # 95% confidence interval
    critical_value <- qt(1 - alpha/2, df = model_adjusted$df.residual)
    robust_coefficients$CI_Lower <- robust_coefficients$Coefficient - critical_value * robust_coefficients$Std_Error
    robust_coefficients$CI_Upper <- robust_coefficients$Coefficient + critical_value * robust_coefficients$Std_Error
    
    # Get R-squared from the original model
    glance_adjusted <- glance(model_adjusted)
    
    # Add model, outcome, and hormone information
    robust_coefficients$Model <- "Robust Adjusted"
    robust_coefficients$Outcome <- outcome
    robust_coefficients$Hormone <- hormone
    robust_coefficients$R_squared <- glance_adjusted$r.squared
    
    # Reorder columns
    robust_coefficients <- robust_coefficients[, c("Model", "Outcome", "Hormone", "Term", "Coefficient", "Std_Error",
                                                   "CI_Lower", "CI_Upper", "p_value", "R_squared")]
    
    # Append to the robust adjusted results data frame
    robust_adjusted_results <- rbind(robust_adjusted_results, robust_coefficients)
    
  }  # End of outcome loop
  
  # Append to the combined robust adjusted results data frame
  all_robust_adjusted_results <- rbind(all_robust_adjusted_results, robust_adjusted_results)
  
}  # End of hormone loop

# Save the combined robust adjusted results to a CSV file
write.csv(all_robust_adjusted_results, file = "Hormone_associations/All/Bodycomp_Robust_Adjusted_results.csv", row.names = FALSE)
```

## 5. Plot

For this code to produce a plot, results of the linear regression should be presented in an excel file. Each row represents a variable with the variable name in column A. The beta-coefficient and p-value of the first independent variable (e.g. E2) should be presented in columns B and C respectively. The beta-coefficients and p-values of the remaining hormones should follow in the same pattern. 

```{r}
# Make sure to read your data properly
final_df <- read_xlsx("Regression_matrix_data.xlsx")
print(final_df)

# Convert variable names to rownames (these will be the labels on the plot)
corrPlot <- column_to_rownames(final_df, var="Variable")

# Extract correlation and p-value matrices (column headings)
corr <- as.matrix(corrPlot[, c("Beta_E2", "Beta_FEI", "Beta_prog", "Beta_TT", "Beta_FAI", "Beta_TE2")])
p <- as.matrix(corrPlot[, c("p_E2", "p_FEI", "p_prog", "p_TT", "p_FAI", "p_TE2")])

# Rename columns 
colnames(corr) <- c("E2", "FEI", "Progesterone", "TT", "FAI", "TT/E2 Ratio")
colnames(p) <- c("E2", "FEI", "Progesterone", "TT", "FAI", "TT/E2 Ratio")

tiff("Hormone_associations/Regression_matrix.tiff", width = 10, height = 15, units = "in", res = 200)

my_colors <- colorRampPalette(c("#313695", "#74add1", "#f7f7f7", "#f46d43", "#a50026"))(100)

c1 <- corrplot(corr,
              insig = 'label_sig',
             order = "original",
             p.mat = p,
             sig.level = 0.05,
             method = "color",
             type = "full",
             tl.cex = 1.8,
             tl.col = "white",
             tl.srt = 45,
             col = my_colors,  # Use custom color palette
             cl.pos = 'r',
             addgrid.col = "gray",
             cl.cex = 1.5,
             cl.ratio = 0.5,
             mar = c(6, 6, 8, 0),
             zlim = c(-1, 1))  # Force scale to be between -0.6 and 0.6

# Left-align the hormone titles using adj=0
text(1:6, 25.8,
     expression("E2", "FEI", "Progesterone", "TT", "FAI", "TT/E2 Ratio"),
     cex = 1.5, font = 2,
     srt = 60, adj = 0)  # Added adj=0 for left alignment

positions <- seq(25, 1, length.out = 25)
mtext(side = 2, line = -25, at = positions,
      text = rownames(corr),
      cex = 1.3, font = 1, adj = 1, las = 2)
dev.off()
```

